---
title: "Report Gapminder - R Introduction Course - CODATA-RDA SS"
author: "Caroline Franco"
date: "August 8, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we will present a report concerning an investigation motivates by the following question:

**Is the *per capita* GDP of a country correlated with its life expectancy?**

We will use the information contained in the *gapminder dataset* of *R*:

```{r gapminder data}
#library(tidyverse)

if (!file.exists("../data/gapminder.csv")) {
  download.file(url = "http://tiny.cc/SWCgapminder",
              destfile = "../data/gapminder.csv")
}
gapminder <- read.csv("../data/gapminder.csv")
head(gapminder)
unique(gapminder$continent)
```

For simplicity, let's restrict ourselves to European countries.

# Life expectancy in American countries over the years

```{r lifeExp}
gapminder %>% 
  filter(continent == 'Americas')  %>% 
  select(country, year, lifeExp) %>% 
  ggplot(aes(x=year, y=lifeExp, colour=country)) +
    geom_path() +
    labs(x="year", y="life expectancy")
```

# *Per capita* GDP in American countries over the years

```{r gdp}
gapminder %>% 
  filter(continent == 'Americas')  %>% 
  select(country, year, gdpPercap) %>% 
  ggplot(aes(x=year, y=gdpPercap, colour=country)) +
    geom_path() +
    labs(x="year", y="GDP per capita")
```

# Correlation between the two sets o data



```{r Correlation}
library(dplyr)
library(plotly)

countries_Americas <-unique(gapminder %>% filter(continent == 'Americas')  %>%  select(country))
countries_Americas %>% as.character()
l_contr <- nrow(countries_Americas)
countries_Am_char <- (as.character(countries_Americas[,1]))

correlations <- c()

for(which_country in 1:l_contr){
  gdp <- gapminder %>% 
    filter(continent == 'Americas', country == countries_Am_char[which_country])  %>% 
    select(gdpPercap, year)
  
  lex <- gapminder %>% 
    filter(continent == 'Americas', country == countries_Am_char[which_country])  %>% 
    select(lifeExp, year)
  
  cor(gdp$gdpPercap, lex$lifeExp)
  
  correlations <- rbind(correlations, cbind(countries_Am_char[which_country],cor(gdp$gdpPercap, lex$lifeExp)))
  head(correlations)
}

corr <- as.data.frame( cbind(as.numeric(1:l_contr), correlations) )
corr2 <- corr[order(corr$V3),]
corr3 <- data.frame()
corr3 <- cbind(as.character(corr2[,2]), as.numeric(corr2[,3]))

ggplot(corr3, aes(x=corr[,1], y=corr[,2])) +
    geom_point() +
    labs(x="country", y="correlation") +
    scale_y_continuous(breaks = c(-0.25, 0, 0.25, 0.5, 0.75, 1))

```
```{r pressure, echo=FALSE}
plot(pressure)
```
